Extended Data

### 2021-10-15 doc/manuscript/Dow_manuscript_Nature.Rmd
discussion
- cite extended tree-ring figure (issue #79 )
- adjust text on previous spring temperatures (issue #73 )

##figures
EDF_map = "Extended Data Figure 1" 
EDF_growth_curves_all = "Extended Data Figure 2"  
EDF_climwin = "Extended Data Figure 3"
EDF_pheno_all = "Extended Data Figures 4-6"
EDF_pheno_Tmax_climwin = "Extended Data Figure 4"
EDF_pheno_Tmax_AprilMay = "Extended Data Figure 5"
EDF_pheno_Tmin_AprilMay = "Extended Data Figure 6"
EDF_tree_rings_TMIN = "Extended Data Figure 7"

@@

<!-- it remains unclear where extra carbon goes (1 par, maybe 2)-->
Our finding that interannual variation in woody growth is more strongly linked to conditions during the peak growing season (Fig. `r tree_rings`) than to growing season length aligns with parallel findings for NEE [@zhou_explaining_2016; @fu_maximum_2019].
However, there is also a disconnect with findings that NEE increases -- at least modestly -- with spring temperatures[@keenan_net_2014] or with the length of the carbon uptake period[@zhou_explaining_2016; @fu_maximum_2019].
We show that this extra C sequestration in years with warm springs does not substantially augment woody growth, raising the question of where it goes.

There are two main possibilities, which hold contrasting implications for the response of forest C balance to rising spring temperatures.
First, C that is fixed in late summer or fall may be used for woody growth that is not apparent from diameter measurements in the current year: either used for thickening of cell walls, a process that lags behind stem expansion[@cuny_woody_2015], or saved within trees as non-structural carbohydrates and used towards growth the following year[@trumbore_quantifying_2002;@richardson_seasonal_2013].
In fact, [extension of our tree-ring analysis](https://github.com/EcoClimLab/growth_phenology/issues/57) showed a generally positive, although often non-significant, correlation between ring-width and previous year growing season $T_{max}$ in RP trees, indicating this may be a possibility. *very little correlation between April $T_{max}$ and growth the following year (5 of 142 RP and 3 of 66 DP species-site combinations), refuting the latter possibility*.
We also cannot rule out the possibility that warm springs promote formation of more carbon-dense cell walls in the current of following year (**coauthors, do any of you know literature related to this?**).

*There may be a shift towards greater belowground allocation[@zohner_how_2021].*

A second possibility is that any additional C fixed during years with warm springs that is not rapidly respired may be allocated, potentially in the following growing season, to plant functions other than stem growth, including respiration, reproduction, foliage, fine roots, or root exudates[@trumbore_quantifying_2002]. 
Stem growth is more sensitive to water deficit than is photosynthesis[@muller_water_2011]. [This seems to promote the investment of sequestered carbon into root growth in times of drought stress. Zoehner et al 2021[@zohner_how_2021] found that the proportion of root-to-stem growth increased under spring warming, suggesting that increased ecosystem sequestration resulting from spring warming may be promoting the growth of below-ground biomass. ~~However, this study was done under tightly controlled conditions within potted plants and does not represent the environmental conditions of forests.~~](https://github.com/EcoClimLab/growth_phenology/issues/72)
Most of this carbon would have a relatively low residence time in the ecosystem[@trumbore_quantifying_2002], although a small fraction would be incorporated into soil organic carbon pools *with residence times rivaling that of wood [REFS]*.
Notably, turnover of extra C fixed during years with warm springs would often occur with sufficient temporal lag that it would not be captured in interannual variation in NEE, implying that the long-term effect of lengthened growing season on NEE may not be captured in studies that consider time frames of a year or less[@zhou_explaining_2016; @fu_maximum_2019].
Studies examining long-term trends in growing season length and ecosystem C uptake[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014] -- as opposed to their interannual variation -- within or including the temperate deciduous biome showed increasing trends in both variables, suggesting that the C not allocated to current year growth may have a *fairly long* residence time within the ecosystem; however, the amount and fate of this carbon remains unknown.
*BUT SEE @piao_net_2008!! and @barichivich_large-scale_2013 - these are focused on higher latitudes, but shows greater release of C in the fall!*

>>>>

##figures
EDF_map = "Extended Data Figure 1" 
EDF_growth_curves_all = "Extended Data Figure 2"  
EDF_climwin = "Extended Data Figure 3"
EDF_pheno_all = "Extended Data Figures 4-6"
EDF_pheno_Tmax_climwin = "Extended Data Figure 4"
EDF_pheno_Tmax_AprilMay = "Extended Data Figure 5"
EDF_tree_rings_TMAX_extended = "Extended Data Figure 6"
EDF_tree_rings_TMIN = "Extended Data Figure 7"

@@

<!-- it remains unclear where extra carbon goes (1 par, maybe 2)-->
Our finding that interannual variation in woody growth is more strongly linked to conditions during the peak growing season (Fig. `r tree_rings`) than to growing season length aligns with parallel findings for NEE [@zhou_explaining_2016; @fu_maximum_2019].
However, there is also a disconnect with findings that NEE increases -- at least modestly -- with spring temperatures[@keenan_net_2014] or with the length of the carbon uptake period[@zhou_explaining_2016; @fu_maximum_2019].
We show that this extra C sequestration in years with warm springs does not substantially augment woody growth, raising the question of where it goes.
There are two main possibilities, which hold contrasting implications for the response of forest C balance to rising spring temperatures.

First, extra C production in years with warm springs may be allocated to woody growth without affecting diameter growth in the current year. 
It is possible that extra carbon is allocated to cell wall thickening, a process that lags behind stem expansion[@cuny_woody_2015], resulting in formation of more carbon-dense cell walls in years with warm springs (**coauthors, do any of you know studies that have looked at this?**).
Extra carbon could also be saved within trees as non-structural carbohydrates and used towards growth the following year[@trumbore_quantifying_2002;@richardson_seasonal_2013]. 
Extension of our tree-ring analysis revealed little correlation between April $T_{max}$ and growth the following year (sig. pos. correlations for 5/142 RP and 3/66 DP species-site combinations, Fig. `r EDF_tree_rings_TMAX_extended`), although predominantly positive (non-significant) correlations in RP species suggests that this dynamic may weakly influence their annual growth.
Finally, extra C may be allocated to growth of coarse roots[@zohner_how_2021] or branches<!-- REF?--> and therefore be unapparent in the basal portions of trunks.
Warm springs could increase woody productivity through any of these mechanisms, yet none are likely to provide substantial, sustained C sinks under warming spring temperatures.

A second possibility is that any additional C fixed during years with warm springs may be allocated, potentially in the following growing season, to plant functions other than stem growth, including respiration, reproduction, foliage, fine roots, or root exudates[@trumbore_quantifying_2002]. 
Stem growth is more sensitive to the water deficits that may be associated with high temperatures than is photosynthesis[@muller_water_2011], and hot or dry conditions have been observed to shift allocation towards roots[@zohner_how_2021], reproduction[**REFS**], or respiration[**REFS**].
Most of this carbon would have a relatively low residence time in the ecosystem[@trumbore_quantifying_2002], although a small fraction would be incorporated into soil organic carbon pools *with residence times rivaling that of wood [REFS]*.
Notably, turnover of extra C fixed during years with warm springs would often occur with sufficient temporal lag that it would not be captured in interannual variation in NEE, implying that the long-term effect of lengthened growing season on NEE may not be captured in studies that consider time frames of a year or less[@zhou_explaining_2016; @fu_maximum_2019].
Studies examining long-term trends in growing season length and ecosystem C uptake[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014] -- as opposed to their interannual variation -- within or including the temperate deciduous biome showed increasing trends in both variables, suggesting that the C not allocated to current year growth may have a *fairly long* residence time within the ecosystem; however, the amount and fate of this carbon remains unknown.
*BUT SEE @piao_net_2008!! and @barichivich_large-scale_2013 - these are focused on higher latitudes, but shows greater release of C in the fall!*
